---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# List comprehensions

<!-- badges: start -->
[![R build status](https://github.com/dirkschumacher/listcomp/workflows/R-CMD-check/badge.svg)](https://github.com/dirkschumacher/listcomp/actions)
[![Lifecycle: maturing](https://img.shields.io/badge/lifecycle-maturing-blue.svg)](https://www.tidyverse.org/lifecycle/#maturing)
[![CRAN status](https://www.r-pkg.org/badges/version/listcomp)](https://CRAN.R-project.org/package=listcomp)
<!-- badges: end -->

The package implements [list comprehensions](https://en.wikipedia.org/wiki/List_comprehension) as purely syntactic sugar with a minor runtime overhead. It constructs nested for-loops and executes the byte-compiled loops to collect the results.

## Installation

``` r
remotes::install_github("dirkschumacher/listcomp")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(listcomp)
head(gen_list(c(x, y), x = 1:100, y = 1:100, z = 1:100, x < 5, y < 5, z == x + y))
```

```{r}
gen_list(c(x, y), x = 1:10, y = x:5, x < 2)
```

This is how the code looks like:
```{r}
lst_verbose <- function(expr, ...) {
  deparse(listcomp:::translate(rlang::enquo(expr), rlang::enquos(...)))
}
lst_verbose(c(x, y), x = 1:10, y = x:5, x < 2)
```

You can also burn in external variables
```{r}
z <- 10
gen_list(c(x, y), x = 1:!!z, y = x:5, x < 2)
```

It also supports parallel iteration by passing a list of named sequences

```{r}
gen_list(c(i, j, k), list(i = 1:10, j = 1:10), k = 1:5, i < 3, k < 3)
```

The code then looks like this:

```{r}
lst_verbose(c(i, j, k), list(i = 1:10, j = 1:10), k = 1:5, i < 3, k < 3)
```

It is quite fast, but the order of filter conditions also greatly determines the execution time.
Sometimes, ahead of time compiling is slower than running it right away.

```{r}
bench::mark(
  a = gen_list(c(x, y), x = 1:100, y = 1:100, z = 1:100, x < 5, y < 5, z == x + y),
  b = gen_list(c(x, y), x = 1:100, x < 5, y = 1:100, y < 5, z = 1:100, z == x + y),
  c = gen_list(c(x, y), x = 1:100, y = 1:100, z = 1:100, x < 5, y < 5, z == x + y, .compile = FALSE),
  d = gen_list(c(x, y), x = 1:100, x < 5, y = 1:100, y < 5, z = 1:100, z == x + y, .compile = FALSE)
)
```

How slow is it compared to a for loop and lapply for a very simple example?

```{r}
bench::mark(
  a = gen_list(x * 2, x = 1:1000, x**2 < 100),
  b = gen_list(x * 2, x = 1:1000, x**2 < 100, .compile = FALSE),
  c = lapply(Filter(function(x) x**2 < 100, 1:1000), function(x) x * 2),
  d = {
    res <- list()
    for (x in 1:1000) {
      if (x**2 >= 100) next
      res[[length(res) + 1]] <- x * 2
    }
    res
  }, 
  time_unit = "ms"
)
```

# Prior art

* [lc](https://github.com/mailund/lc) Uses a similiar syntax as `complist`
* [comprehenr](https://github.com/gdemin/comprehenr) Uses a similiar code generation approach as `complist` but with a different syntax.

